---
title: "Exosomes raw data Quality control"
author: "David Cáceres"
date: "30/1/2023"
output:
  pdf_document: 
    toc: yes
    toc_depth: 2
    fig_height: 3
    fig_width: 6
  html_document:
    fig_caption: yes
    toc: yes
    toc_depth: 2
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir ="~/Exosomas/QC/Data/",
                      tidy.opts=list(width.cutof=30),tidy=TRUE,
                      comment = NULL, cache = TRUE,dev = 'pdf')


```

```{r, echo=FALSE,message=FALSE}
library(ggplot2)
library(DESeq2)
library(edgeR)
library(patchwork)
library(ggsci)
library(cowplot)
library(data.table)
library(tibble)
library(factoextra)
library(base)
library(tidyverse)
library(gridExtra)
library(plyr)
library(swamp)
```


## Data read

```{r}
setwd("~/Exosomas/QC/Data/")
tipos_exo<-read.table(file="readsPerc.tsv", header = TRUE, sep = "\t")
brainq_reads<-read.table(file="brain_reads.tsv", header = TRUE, sep = "\t")
brainq_short<-read.table(file="brainq_short.tsv", header = TRUE, sep = "\t")
brainq_ushort<-read.table(file="brainq_ushort.tsv", header = TRUE, sep = "\t")
short<-read.table(file="shortReadsPerc.tsv", header = TRUE, sep = "\t")
ushort<-read.table(file="ultrashort.tsv", header = TRUE, sep = "\t")
ref.genome<-read.table(file="ref.genome.tsv", header = TRUE, sep = "\t")
unmapped<-read.table(file="unmapped.tsv", header = TRUE, sep = "\t")
contaminacion<-read.table(file="Contaminacion.tsv", header = TRUE, sep = "\t")
brainq_contaminacion<-read.table(file="brainq_contaminacion.tsv", header = TRUE,sep = "\t")
reads_total<-read.table(file="Total_reads.tsv", header = TRUE, sep = "\t")

```


# Reads distribution by batch with brain replicate control


## Effective reads percent distribution 

We used miRNAQC to get quality reports and datasets used to pre-process the raw data and check it quality.
arn.ugr.es/mirnaqc/

Initially we plotted the effective reads percent vs the sample distribution by batch. Also plotted the percernt of short reads and not adapter found. As reference we plotted the brain replicate data.


```{r warning=FALSE}

# Adding shortnames

values <- c("A","B","C","D","E","F","G", "H")
tipos_exo$batch<-as.factor(tipos_exo$batch)
tipos_exo$batch_shortname <- values[tipos_exo$batch]

# Percent of effective reads

tipos_exo <- tipos_exo |> 
    mutate(sample1 = reorder(sample, -ifelse(!X %in% "readsPerc", 0, value), 
                             FUN = sum))

p1 <- ggplot(tipos_exo,aes(x = sample1, y = value, fill = factor(X, levels=c("readsAdapterNotFoundPerc","shortReads","readsPerc")), width=1.05)) +
    geom_bar( stat = "identity", position = position_stack()) +
    scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10)) +
    labs(x = NULL, y= "% Reads" , title="Efective reads% Distribution", 
         fill="Quality Control") +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.length.x = unit(0, "pt")) +
    theme(
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.background = element_rect(fill='transparent'),
        legend.box.background = element_rect(fill='transparent'),
        plot.title = element_text(face="bold")
    ) + scale_fill_simpsons(labels=c("Adapter not found","Short Reads%","Reads%"
                                     ))
tipos_exo1 <- tipos_exo |>
    distinct(sample, sample1, batch_shortname)

p_axis <- ggplot(tipos_exo1, aes(x = sample1, y = factor(1), 
                                 fill = batch_shortname)) +
    geom_tile(width = 1) +
    theme_void() +
    theme(axis.title.x = element_text()) + theme(legend.position = "none")  +
    labs(x="Batch Annotation", fill="Batch") 

p1_q<-p1 / p_axis + plot_layout(heights = c(8,1))  + scale_fill_igv()

# Brain

p_reads_brain<-ggplot(brainq_reads,aes(x = batch, y = value1, 
                                       fill = factor(variable1, levels=c("readsAdapterNotFoundPerc","shortReads","readsPerc")))) +
geom_bar( stat = "identity") +
scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10)) +
labs(x = NULL, y= "% Reads" , title="Brain Reads%") +
theme(axis.text.x = element_blank(),
axis.ticks.length.x = unit(0, "pt")) + scale_fill_simpsons()+
theme(
panel.background = element_rect(fill='transparent'),
plot.background = element_rect(fill='transparent', color=NA),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
legend.background = element_rect(fill='transparent'),
legend.box.background = element_rect(fill='transparent'),
plot.title = element_text(face="bold")
) + theme(legend.position = "none")
p_axis_reads_brain <- ggplot(brainq_reads, aes(x =batch, y = factor(1), 
                                               fill = batch_shortname)) +
geom_tile(width = 1) +
theme_void() +
theme(axis.title.x = element_text())  +
geom_text(aes(label = batch_shortname)) +
labs(x="Batch Annotation", fill="Batch Annotation") +
theme(legend.position="none")
p3_qb<-p_reads_brain / p_axis_reads_brain + plot_layout(heights = c(8, 1))+ 
  scale_fill_igv()

pt1<-plot_grid(p1_q,p3_qb, rel_widths = c(1.7, 1.3), labels = c("A","B"))

pt1

```


## Short reads percent. Less than 17 nucleotides

Plot of the short reads obtained from mirnaQC


```{r}
#Short Reads%
short <- short |> 
    mutate(sample1 = reorder(sample, 
                             -ifelse(!X %in% "shortReadsPerc", 0, value), 
                             FUN = sum))

p_short<-ggplot(short,aes(x =sample1, y = value, fill = factor(X, levels=c("readsAdapterNotFoundPerc","reads","shortReadsPerc")), width=1.05)) +
geom_bar( stat = "identity") +
scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10)) +
labs(x = NULL, y= "Short reads%" , title="Short reads% Distribution", 
     fill="Quality Control") +
theme(axis.text.x = element_blank(),
axis.ticks.x = element_blank(),
axis.ticks.length.x = unit(0, "pt")) + 
theme(
panel.background = element_rect(fill='transparent'),
plot.background = element_rect(fill='transparent', color=NA),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
legend.background = element_rect(fill='transparent'),
legend.box.background = element_rect(fill='transparent')) + 
scale_fill_simpsons(labels=c("Adapter not found","Reads%","Short Reads%"))

p_axis_short <- ggplot(short, aes(x =sample1, y = factor(1), fill = batch)) +
geom_tile(width = 1) +
theme_void() +
theme(axis.title.x = element_text()) + theme(legend.position = "none") +
labs(x="Batch Annotation", fill="Batch")
p3q<-p_short / p_axis_short + plot_layout(heights = c(8, 1))  + scale_fill_igv()

# Brain

p_short_brain<-ggplot(brainq_short,aes(x = batch, 
                                       y = value1,fill = factor(variable1, levels=c("readsAdapterNotFoundPerc","reads","shortReadsPerc")))) +
geom_bar( stat = "identity") +
scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10)) +
labs(x = NULL, y= "% Short reads" , title="Brain short reads%", 
     fill="QC") +
theme(axis.text.x = element_blank(),
axis.ticks.x = element_blank(),
axis.ticks.length.x = unit(0, "pt")) + scale_fill_simpsons()+
theme(
panel.background = element_rect(fill='transparent'),
plot.background = element_rect(fill='transparent', color=NA),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
legend.background = element_rect(fill='transparent'),
legend.box.background = element_rect(fill='transparent')
) +
theme(legend.position="none")
p_axis_short_brain <- ggplot(brainq_short, aes(x =batch, y = factor(1), 
                                               fill = batch_shortname)) +
geom_tile(width = 1) +
theme_void() +
theme(axis.title.x = element_text()) +
geom_text(aes(label = batch_shortname)) +
labs(x="Batch Annotation", fill="Batch Annotation") +
theme(legend.position="none") 
p4_qb<-p_short_brain / p_axis_short_brain + plot_layout(heights = c(8, 1)) + 
scale_fill_igv()

pt2<-plot_grid(p3q,p4_qb, rel_widths = c(1.7, 1.3))
pt2
```


## Ultra short reads percent. Less than 15 nucleotides.

```{r}
# Ultra Short%
ushort <- ushort |> 
    mutate(sample1 = reorder(sample, -ifelse(!X %in% "ultraShortReadsPerc", 0,
                                             value), FUN = sum))
p_ushort<-ggplot(ushort,aes(x =sample1, y = value, fill = factor(X, levels=c("readsAdapterNotFoundPerc","reads","ultraShortReadsPerc")), 
                            width=1.05)) +
geom_bar( stat = "identity") +
scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10)) +
labs(x = NULL, y= "Ultra Short reads%" , 
     title="Ultra Short Reads% Distribution", fill="Quality Control") +
theme(axis.text.x = element_blank(),
axis.ticks.x = element_blank(),
axis.ticks.length.x = unit(0, "pt")) + 
theme(
panel.background = element_rect(fill='transparent'),
plot.background = element_rect(fill='transparent', color=NA),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
legend.background = element_rect(fill='transparent'),
legend.box.background = element_rect(fill='transparent'),
plot.title = element_text(face="bold")
) + scale_fill_simpsons(labels=c("Adapter not found","Reads%",
                                 "Ultra short Reads%"))

p_axis_ushort <- ggplot(ushort, aes(x =sample1, y = factor(1), fill = batch)) +
geom_tile(width = 1) +
theme_void() +
theme(axis.title.x = element_text()) + theme(legend.position = "none") +
labs(x="Batch Annotation", fill="Batch")
p5_q<-p_ushort / p_axis_ushort + plot_layout(heights = c(8, 1)) +
  scale_fill_igv()

# Brain

p_ushort_brain<-ggplot(brainq_ushort,aes(x = batch, 
                                         y = value1,fill = factor(variable1, levels=c("readsAdapterNotFoundPerc","reads","ultraShortReadsPerc")))) +
geom_bar( stat = "identity") +
scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10)) +
labs(x = NULL, y= "% Ultra Short reads" , 
     title="Brain Ultra short reads%", fill="QC") +
theme(axis.text.x = element_blank(),
axis.ticks.x = element_blank(),
axis.ticks.length.x = unit(0, "pt")) + scale_fill_simpsons()+
theme(
panel.background = element_rect(fill='transparent'),
plot.background = element_rect(fill='transparent', color=NA),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
legend.background = element_rect(fill='transparent'),
legend.box.background = element_rect(fill='transparent'),
plot.title = element_text(face="bold")
) +
theme(legend.position="none")

p_axis_ushort_brain <- ggplot(brainq_ushort, aes(x =batch, y = factor(1),
                                                 fill = batch_shortname)) +
geom_tile(width = 1) +
theme_void() +
theme(axis.title.x = element_text()) +
geom_text(aes(label = batch_shortname)) +
labs(x="Batch Annotation", fill="Batch Annotation") +
theme(legend.position="none") 
p5_qb<-p_ushort_brain / p_axis_ushort_brain + plot_layout(heights = c(8, 1))  +
  scale_fill_igv()

pt3<-plot_grid(p5_q,p5_qb, rel_widths = c(1.7, 1.3))
pt3
```


## Reference genome contamination percent.

```{r}
contaminacion$batch<-as.factor(contaminacion$batch)
contaminacion$batch_shortname <- values[contaminacion$batch]

# Referece genome contamination

contaminacion <- contaminacion |> 
    mutate(sample1 = reorder(sample, -ifelse(!variable %in% "% ref.genome", 
                                             0, value), FUN = sum))
p_ref.genome<-ggplot(contaminacion,
                     aes(x =sample1, y = value, 
                                       fill = factor(variable, 
                                                     levels=c("virus",
                                                              "bacteria",
                                                              "unmapped",
                                                              "% ref.genome")), 
                         width=1.05)) +
geom_bar( stat = "identity") +
scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10)) +
labs(x = NULL, y= "Reads%" , title="Mapped Reads% Distribution", 
     fill="Quality Control") +
theme(axis.text.x = element_blank(),
axis.ticks.x = element_blank(),
axis.ticks.length.x = unit(0, "pt")) + 
theme(
panel.background = element_rect(fill='transparent'),
plot.background = element_rect(fill='transparent', color=NA),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
legend.background = element_rect(fill='transparent'),
legend.box.background = element_rect(fill='transparent')
) + scale_fill_simpsons(labels=c("Virus","Bacteria","Unmapped","Mapped Reads%"))

contaminacion1 <- contaminacion |>
    distinct(sample, sample1, batch_shortname)
p_axis_ref.genome <- ggplot(contaminacion1, aes(x =sample1, y = factor(1), 
                                                fill = batch_shortname)) +
geom_tile(width = 1) +
theme_void() +
theme(axis.title.x = element_text()) + theme(legend.position = "none") +
labs(x="Batch Annotation", fill="Batch")
p6_q<-p_ref.genome / p_axis_ref.genome + plot_layout(heights = c(8, 1))  +
  scale_fill_igv()

# Brain

brainq_contaminacion$batch<-as.factor(brainq_contaminacion$batch)
brainq_contaminacion$batch_shortname <- values[brainq_contaminacion$batch]

p_ref.genome_brain<-ggplot(brainq_contaminacion,
                           aes(x = batch, y = value,fill = factor(variable, levels=c("virus","bacteria","unmapped","refSpecies")))) +
geom_bar( stat = "identity") +
scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10)) +
labs(x = NULL, y= "% Reads" , title="Mapped Reads%", fill="QC") +
theme(axis.text.x = element_blank(),
axis.ticks.x = element_blank(),
axis.ticks.length.x = unit(0, "pt")) + scale_fill_simpsons()+
theme(
panel.background = element_rect(fill='transparent'),
plot.background = element_rect(fill='transparent', color=NA),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
legend.background = element_rect(fill='transparent'),
legend.box.background = element_rect(fill='transparent')
) +
theme(legend.position="none")


p_axis_ref.genome_brain <- ggplot(brainq_contaminacion, 
                                  aes(x =batch, y = factor(1), 
                                      fill = batch_shortname)) +
geom_tile(width = 1) +
theme_void() +
theme(axis.title.x = element_text()) +
geom_text(aes(label = batch_shortname)) +
labs(x="Batch Annotation", fill="Batch Annotation") +
theme(legend.position="none") 
p6_qb<-p_ref.genome_brain / p_axis_ref.genome_brain+
  plot_layout(heights = c(8, 1))  + scale_fill_igv()

plot_grid(p6_q,p6_qb, rel_widths = c(2, 1))

```


## Unmapped reads percent.


```{r}
# Unmapped reads%

contaminacion <- contaminacion |> 
    mutate(sample1 = reorder(sample, -ifelse(!variable %in% "unmapped", 0, 
                                             value), FUN = sum))

p_unmapped<-ggplot(contaminacion,aes(x =sample1, y = value, 
                                     fill = factor(variable, 
                                                   levels=c("virus",
                                                            "bacteria",
                                                            "% ref.genome",
                                                            "unmapped")), 
                                     width=1.05)) +
geom_bar( stat = "identity") +
scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10)) +
labs(x = NULL, y= "Reads%" , title="Unmapped Reads% Distribution", 
     fill="Quality Control") +
theme(axis.text.x = element_blank(),
axis.ticks.x = element_blank(),
axis.ticks.length.x = unit(0, "pt")) + 
theme(
panel.background = element_rect(fill='transparent'),
plot.background = element_rect(fill='transparent', color=NA),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
legend.background = element_rect(fill='transparent'),
legend.box.background = element_rect(fill='transparent')
) + scale_fill_simpsons(labels=c("Virus","Bacteria","Mapped Reads%","Unmapped"))


contaminacion1 <- contaminacion |>
    distinct(sample, sample1, batch_shortname)

p_axis_unmapped <- ggplot(contaminacion1, aes(x =sample1, y = factor(1),
                                              fill = batch_shortname)) +
geom_tile(width = 1) +
theme_void() +
theme(axis.title.x = element_text()) + theme(legend.position = "none") +
labs(x="Batch Annotation", fill="Batch")
p7_q<-p_unmapped / p_axis_unmapped + plot_layout(heights = c(8, 1)) +
  scale_fill_igv()

# Brain

p_unmapped_brain<-ggplot(brainq_contaminacion,aes(x = batch, y = value,
                                                  fill = factor(variable, levels=c("virus","bacteria","unmapped","refSpecies")))) +
geom_bar( stat = "identity") +
scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10)) +
labs(x = NULL, y= "% Reads" , title="Unmapped Reads%", fill="QC") +
theme(axis.text.x = element_blank(),
axis.ticks.x = element_blank(),
axis.ticks.length.x = unit(0, "pt")) + scale_fill_simpsons()+
theme(
panel.background = element_rect(fill='transparent'),
plot.background = element_rect(fill='transparent', color=NA),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
legend.background = element_rect(fill='transparent'),
legend.box.background = element_rect(fill='transparent')
) +
theme(legend.position="none")

p_axis_unmapped_brain <- ggplot(brainq_contaminacion, 
                                aes(x =batch, y = factor(1), 
                                    fill = batch_shortname)) +
geom_tile(width = 1) +
theme_void() +
theme(axis.title.x = element_text()) +
geom_text(aes(label = batch_shortname)) +
labs(x="Batch Annotation", fill="Batch Annotation") +
theme(legend.position="none") 
p7_qb<-p_unmapped_brain/p_axis_unmapped_brain + plot_layout(heights = c(8, 1)) +
  scale_fill_igv()

plot_grid(p7_q,p7_qb, rel_widths = c(2, 1))

```


## Bacteria reads percent. Contamination.

```{r}
# Bacteria

contaminacion <- contaminacion |> 
    mutate(sample1 = reorder(sample, -ifelse(!variable %in% "bacteria", 
                                             0, value), FUN = sum))

p_bacteria<-ggplot(contaminacion,
                   aes(x =sample1, y = value,
                       fill = factor(variable, 
                                     levels=c("% ref.genome","unmapped","virus",
                                              "bacteria")), width=1.05)) +
geom_bar( stat = "identity") +
scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10)) +
labs(x = NULL, y= "Reads%" , title="Bacteria Reads%", fill="Quality Control") +
theme(axis.text.x = element_blank(),
axis.ticks.x = element_blank(),
axis.ticks.length.x = unit(0, "pt")) + 
theme(
panel.background = element_rect(fill='transparent'),
plot.background = element_rect(fill='transparent', color=NA),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
legend.background = element_rect(fill='transparent'),
legend.box.background = element_rect(fill='transparent')
) + scale_fill_simpsons(labels=c("Mapped Reads%","Unmapped","Virus","Bacteria"))

contaminacion1 <- contaminacion |>
    distinct(sample, sample1, batch_shortname)

p_axis_bacteria <- ggplot(contaminacion1, aes(x =sample1, y = factor(1),
                                              fill = batch_shortname)) +
geom_tile(width = 1) +
theme_void() +
theme(axis.title.x = element_text()) + theme(legend.position = "none") +
labs(x="Batch Annotation", fill="Batch")
p8_q<-p_bacteria / p_axis_bacteria + plot_layout(heights = c(8, 1)) +
  scale_fill_igv()

# Brain

p_bacteria_brain<-ggplot(brainq_contaminacion,aes(x = batch, y = value,
                                                  fill = factor(variable, levels=c("unmapped","refSpecies","virus","bacteria")))) +
geom_bar( stat = "identity") +
scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10)) +
labs(x = NULL, y= "% Reads" , title="Brain Bacteria Reads%", fill="QC") +
theme(axis.text.x = element_blank(),
axis.ticks.x = element_blank(),
axis.ticks.length.x = unit(0, "pt")) + scale_fill_simpsons()+
theme(
panel.background = element_rect(fill='transparent'),
plot.background = element_rect(fill='transparent', color=NA),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
legend.background = element_rect(fill='transparent'),
legend.box.background = element_rect(fill='transparent')
) +
theme(legend.position="none")

p_axis_bacteria_brain <- ggplot(brainq_contaminacion, 
                                aes(x =batch, y = factor(1), 
                                    fill = batch_shortname)) +
geom_tile(width = 1) +
theme_void() +
theme(axis.title.x = element_text()) +
geom_text(aes(label = batch_shortname)) +
labs(x="Batch Annotation", fill="Batch Annotation") +
theme(legend.position="none") 

p8_qb<-p_bacteria_brain/p_axis_bacteria_brain + plot_layout(heights = c(8, 1))+
  scale_fill_igv()


pt8<-plot_grid(p8_q,p8_qb, rel_widths = c(1.7, 1.3))
pt8
```


## Virus reads percent. Contamination

```{r}
# Virus

contaminacion <- contaminacion |> 
    mutate(sample1 = reorder(sample, -ifelse(!variable %in% "virus", 0, value), 
                             FUN = sum))
p_virus<-ggplot(contaminacion,
                aes(x =sample1, y = value, fill = factor(variable, levels=c("% ref.genome","unmapped","bacteria","virus")), width=1.05)) +
geom_bar( stat = "identity") +
scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10)) +
labs(x = NULL, y= "Reads%" , title="Virus Reads%", fill="Quality Control") +
theme(axis.text.x = element_blank(),
axis.ticks.x = element_blank(),
axis.ticks.length.x = unit(0, "pt")) + 
theme(
panel.background = element_rect(fill='transparent'),
plot.background = element_rect(fill='transparent', color=NA),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
legend.background = element_rect(fill='transparent'),
legend.box.background = element_rect(fill='transparent')
) + scale_fill_simpsons(labels=c("Mapped Reads%","Unmapped","Bacteria","Virus"))

contaminacion1 <- contaminacion |>
    distinct(sample, sample1, batch_shortname)

p_axis_virus <- ggplot(contaminacion1, aes(x =sample1, y = factor(1), 
                                           fill = batch_shortname)) +
geom_tile(width = 1) +
theme_void() +
theme(axis.title.x = element_text()) + theme(legend.position = "none") +
labs(x="Batch Annotation", fill="Batch")

p9_q<-p_virus / p_axis_virus + plot_layout(heights = c(8, 1)) +
  scale_fill_igv()

# Brain

p_virus_brain<-ggplot(brainq_contaminacion,aes(x = batch, y = value,
                                               fill = factor(variable, levels=c("unmapped","refSpecies","bacteria","virus")))) +
geom_bar( stat = "identity") +
scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10)) +
labs(x = NULL, y= "% Reads" , title="Brain Virus Reads%", fill="QC") +
theme(axis.text.x = element_blank(),
axis.ticks.x = element_blank(),
axis.ticks.length.x = unit(0, "pt")) + scale_fill_simpsons()+
theme(
panel.background = element_rect(fill='transparent'),
plot.background = element_rect(fill='transparent', color=NA),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
legend.background = element_rect(fill='transparent'),
legend.box.background = element_rect(fill='transparent')
) +
theme(legend.position="none")

p_axis_virus_brain <- ggplot(brainq_contaminacion, 
                             aes(x =batch, y = factor(1), 
                                 fill = batch_shortname)) +
geom_tile(width = 1) +
theme_void() +
theme(axis.title.x = element_text()) +
geom_text(aes(label = batch_shortname)) +
labs(x="Batch Annotation", fill="Batch Annotation") +
theme(legend.position="none") 
p9_qb<-p_virus_brain / p_axis_virus_brain + plot_layout(heights = c(8, 1))+
  scale_fill_igv()

pt9<-plot_grid(p9_q,p9_qb, rel_widths = c(1.7, 1.3))
pt9
```


## Total reads in absolute value. 

```{r}

reads_total_brain<-reads_total[c(13,42,77,106,141,194,222,256),]

p_reads_t<-ggplot(reads_total,aes(x = reorder(sample, -reads), y = reads,
                                  fill = variable, width=1.05)) +
geom_bar( stat = "identity") +
scale_y_continuous(expand = c(0, 0)) +
labs(x = NULL, y= "Reads" , title="Reads total QC", fill="Quality Control") +
theme(axis.text.x = element_blank(),
axis.ticks.x = element_blank(),
axis.ticks.length.x = unit(0, "pt")) + 
theme(
panel.background = element_rect(fill='transparent'),
plot.background = element_rect(fill='transparent', color=NA),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
legend.background = element_rect(fill='transparent'),
legend.box.background = element_rect(fill='transparent')
) + scale_fill_manual(values = c("darkgrey"),labels=c("Total Reads"))

p_axis_reads_t<- ggplot(reads_total, 
                        aes(x = reorder(sample, -reads),
                            y = factor(1), fill = batch_shortname)) +
geom_tile(width = 1) +
theme_void() +
theme(axis.title.x = element_text(), legend.position=c(1.09,2.2)) +
labs(x="Batch Annotation", fill="Batch Annotation") +
  theme(legend.position = "none") + scale_fill_igv()

p10_q<-p_reads_t / p_axis_reads_t + plot_layout(heights = c(8, 1))  

# Brain

p_reads_total_brain<-ggplot(reads_total_brain,aes(x = sample, y = reads,
                                                  fill = "none")) +
geom_bar( stat = "identity") +
scale_y_continuous(expand = c(0, 0)) +
labs(x = NULL, y= "Reads" , title="Brain Reads total QC") +
theme(axis.text.x = element_blank(),
axis.ticks.x = element_blank(),
axis.ticks.length.x = unit(0, "pt")) +
theme(
panel.background = element_rect(fill='transparent'),
plot.background = element_rect(fill='transparent', color=NA),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
legend.background = element_rect(fill='transparent'),
legend.box.background = element_rect(fill='transparent')
) + theme(legend.position = "none") +
scale_fill_manual(values = c("darkgrey"))

p_axis_total_brain <- ggplot(reads_total_brain, aes(x =sample, y = factor(1), 
                                                    fill = batch_shortname)) +
geom_tile(width = 1) +
geom_text(aes(label = batch_shortname)) +
theme_void() +
theme(axis.title.x = element_text()) +  
labs(x="Batch Annotation") + theme(legend.position = "none") +
scale_fill_igv()

p10_qb<-p_reads_total_brain/ p_axis_total_brain + plot_layout(heights = c(8, 1))  
pt10<-plot_grid(p10_q,p10_qb, rel_widths = c(1.7, 1.3))
pt10
```

# Expression Matrix

## Builiding a expression matrix for microRNA data

We builded a EM from files obtained from srnaToolBox. We did it for each batch, and joined them in a complete expression matrix.

```{r warning=FALSE}

setwd("~/Exosomas/QC/Expresion/Expresion_180322-190603/")
filenames <- list.files(pattern="*.tsv") # Read data
all_files <- lapply(filenames,function(x) { # All files in a list
read.table(file = x,
sep = '\t',
header = TRUE)
})
list3<- lapply(all_files, "[", c(1,3)) # I need these 2 columns
list4<-rbindlist(list3,fill=TRUE) # Binding all DF from the list
list4 <- lapply(list3, function(x) x[!duplicated(x$name),]) # Removing duplicates
matrixA<-rbindlist(list4,fill=TRUE)
matrixA[is.na(matrixA)] = 0
matrixA<-ddply(matrixA,.(name),numcolwise(sum))

write.table(matrixA, 
            file = "~/Exosomas/QC/Expresion/Matrices de Expresion/matrixA.csv", 
            quote = TRUE, sep = ",")

setwd("~/Exosomas/QC/Expresion/Expresion_180626-180628-190531/")
filenames <- list.files(pattern="*.tsv") 
all_files <- lapply(filenames,function(x) { 
read.table(file = x,
sep = '\t',
header = TRUE)
})
list3<- lapply(all_files, "[", c(1,3)) 
list4<-rbindlist(list3,fill=TRUE) 
list4 <- lapply(list3, function(x) x[!duplicated(x$name),])
matrixB<-rbindlist(list4,fill=TRUE)
matrixB[is.na(matrixB)] = 0
matrixB<-ddply(matrixB,.(name),numcolwise(sum))

write.table(matrixB, 
            file = "~/Exosomas/QC/Expresion/Matrices de Expresion/matrixB.csv", 
            quote = TRUE, sep = ",")

setwd("~/Exosomas/QC/Expresion/Expresion_180726-190530/")
filenames <- list.files(pattern="*.tsv") 
all_files <- lapply(filenames,function(x) {
read.table(file = x,
sep = '\t',
header = TRUE)
})
list3<- lapply(all_files, "[", c(1,3)) 
list4<-rbindlist(list3,fill=TRUE) 
list4 <- lapply(list3, function(x) x[!duplicated(x$name),])
matrixC<-rbindlist(list4,fill=TRUE)
matrixC[is.na(matrixC)] = 0
matrixC<-ddply(matrixC,.(name),numcolwise(sum))

write.table(matrixC, 
            file = "~/Exosomas/QC/Expresion/Matrices de Expresion/matrixC.csv",
            quote = TRUE, sep = ",")

setwd("~/Exosomas/QC/Expresion/Expresion_180913-190529/")
filenames <- list.files(pattern="*.tsv") 
all_files <- lapply(filenames,function(x) { 
read.table(file = x,
sep = '\t',
header = TRUE)
})
list3<- lapply(all_files, "[", c(1,3)) 
list4<-rbindlist(list3,fill=TRUE) 
list4 <- lapply(list3, function(x) x[!duplicated(x$name),])
matrixD<-rbindlist(list4,fill=TRUE)
matrixD[is.na(matrixD)] = 0
matrixD<-ddply(matrixD,.(name),numcolwise(sum))

write.table(matrixD, 
            file = "~/Exosomas/QC/Expresion/Matrices de Expresion/matrixD.csv", 
            quote = TRUE, sep = ",")

setwd("~/Exosomas/QC/Expresion/Expresion_180919-190527/")
filenames <- list.files(pattern="*.tsv")
all_files <- lapply(filenames,function(x) { 
read.table(file = x,
sep = '\t',
header = TRUE)
})
list3<- lapply(all_files, "[", c(1,3)) 
list4<-rbindlist(list3,fill=TRUE) 
list4 <- lapply(list3, function(x) x[!duplicated(x$name),])
matrixE<-rbindlist(list4,fill=TRUE)
matrixE[is.na(matrixE)] = 0
matrixE<-ddply(matrixE,.(name),numcolwise(sum))
write.table(matrixE, 
            file = "~/Exosomas/QC/Expresion/Matrices de Expresion/matrixE.csv",
            quote = TRUE, sep = ",")


setwd("~/Exosomas/QC/Expresion/Expresion_181003-190524/")
filenames <- list.files(pattern="*.tsv") 
all_files <- lapply(filenames,function(x) { 
read.table(file = x,
sep = '\t',
header = TRUE)
})
list3<- lapply(all_files, "[", c(1,3)) 
list4<-rbindlist(list3,fill=TRUE) 
list4 <- lapply(list3, function(x) x[!duplicated(x$name),])
matrixF<-rbindlist(list4,fill=TRUE)
matrixF[is.na(matrixF)] = 0
matrixF<-ddply(matrixF,.(name),numcolwise(sum))
write.table(matrixF, 
            file = "~/Exosomas/QC/Expresion/Matrices de Expresion/matrixF.csv", 
            quote = TRUE, sep = ",")


setwd("~/Exosomas/QC/Expresion/Expresion_181004-190521/")
filenames <- list.files(pattern="*.tsv") 
all_files <- lapply(filenames,function(x) {
read.table(file = x,
sep = '\t',
header = TRUE)
})
list3<- lapply(all_files, "[", c(1,3)) 
list4<-rbindlist(list3,fill=TRUE) 
list4 <- lapply(list3, function(x) x[!duplicated(x$name),])
matrixG<-rbindlist(list4,fill=TRUE)
matrixG[is.na(matrixG)] = 0
matrixG<-ddply(matrixG,.(name),numcolwise(sum))
write.table(matrixG, 
            file = "~/Exosomas/QC/Expresion/Matrices de Expresion/matrixG.csv",
            quote = TRUE, sep = ",")


setwd("~/Exosomas/QC/Expresion/Expresion_181023-190319/")
filenames <- list.files(pattern="*.tsv") 
all_files <- lapply(filenames,function(x) {
read.table(file = x,
sep = '\t',
header = TRUE)
})
list3<- lapply(all_files, "[", c(1,3)) 
list4<-rbindlist(list3,fill=TRUE) 
list4 <- lapply(list3, function(x) x[!duplicated(x$name),])
matrixH<-rbindlist(list4,fill=TRUE)
matrixH[is.na(matrixH)] = 0
matrixH<-ddply(matrixH,.(name),numcolwise(sum))
write.table(matrixH, 
            file = "~/Exosomas/QC/Expresion/Matrices de Expresion/matrixH.csv",
            quote = TRUE, sep = ",")


setwd("~/Exosomas/QC/Expresion/Matrices de Expresion/")
filenames <- list.files(pattern="*.csv") 
all_files <- lapply(filenames,function(x) { 
read.delim(file = x,
sep = ',',
header = TRUE,
stringsAsFactors=FALSE)
})
matrixX<-rbindlist(all_files,fill=TRUE)
matrixX<-matrixX %>% mutate_if(is.integer,as.numeric)
matrixX[is.na(matrixX)] = 0
matrixX<-ddply(matrixX,.(name),numcolwise(sum))
matrixX<-matrixX[apply(matrixX[,-1], 1, function(x) !all(x==0)),]

p<-"~/Exosomas/QC/Expresion/Matrices de Expresion/Expression matrix/matrixX.csv"
write.table(matrixX, 
            file = p, quote = TRUE, sep = ",")

file.remove('matrixA.csv','matrixB.csv','matrixC.csv','matrixD.csv',
            'matrixE.csv','matrixF.csv','matrixG.csv','matrixH.csv')
```


```{r}
t<-"~/Exosomas/QC/Expresion/Matrices de Expresion/Expression matrix/matrixX.csv"
Matrix<-read.table(file=t, header = TRUE, sep = ",")

tail(Matrix[,1:4],n=5)
```

# Genome Distribution

## Building a expression matrix for Genome distribution

We builded a new expression matrix to check the average lenght of the microRNA reads.

We are looking for 21-23 nucleotides reads, which is the cannonical lenght for them.


```{r echo=FALSE}
setwd("~/Exosomas/QC/readLen/Expresion_180322-190603/")
filenames <- list.files(pattern="*/*.tsv", recursive=TRUE) # Reading files
all_files <- lapply(filenames,function(x) { # Loading files in a list
read.delim(file = x,
sep = '\t',
header = TRUE)
})
list3<- lapply(all_files, "[", c(1,5))
matrixA<-rbindlist(list3, fill=TRUE) # Binding all dataframes
readLen_RC<-matrixA %>%   # Collapsing data
    group_by(Read.Length..nt.) %>% 
    dplyr::summarise(across(starts_with("x"), ~sum(., na.rm = TRUE)))

readLen_RC_A<-readLen_RC %>% dplyr::select(-contains("Brain"))
names(readLen_RC_A)=stringr::str_sub(names(readLen_RC_A),2)
names(readLen_RC_A)[names(readLen_RC_A) == 'ead.Length..nt.'] <- 'Read_Length_nt'


setwd("~/Exosomas/QC/readLen/Expresion_180626-180628-190531/")
filenames <- list.files(pattern="*/*.tsv", recursive=TRUE) 
all_files <- lapply(filenames,function(x) { 
read.delim(file = x,
sep = '\t',
header = TRUE)
})
list3<- lapply(all_files, "[", c(1,5))
matrixA<-rbindlist(list3, fill=TRUE) 
readLen_RC<-matrixA %>%   
    group_by(Read.Length..nt.) %>% 
    dplyr::summarise(across(starts_with("x"), ~sum(., na.rm = TRUE)))

readLen_RC_B<-readLen_RC %>% dplyr::select(-contains("Brain"))
names(readLen_RC_B)=stringr::str_sub(names(readLen_RC_B),2)
names(readLen_RC_B)[names(readLen_RC_B) == 'ead.Length..nt.'] <- 'Read_Length_nt'


setwd("~/Exosomas/QC/readLen/Expresion_180726-190530/")
filenames <- list.files(pattern="*/*.tsv", recursive=TRUE) 
all_files <- lapply(filenames,function(x) { 
read.delim(file = x,
sep = '\t',
header = TRUE)
})
list3<- lapply(all_files, "[", c(1,5))
matrixA<-rbindlist(list3, fill=TRUE) 
readLen_RC<-matrixA %>%   
    group_by(Read.Length..nt.) %>% 
    dplyr::summarise(across(starts_with("x"), ~sum(., na.rm = TRUE)))

readLen_RC_C<-readLen_RC %>% dplyr::select(-contains("Brain"))
names(readLen_RC_C)=stringr::str_sub(names(readLen_RC_C),2)
names(readLen_RC_C)[names(readLen_RC_C) == 'ead.Length..nt.'] <- 'Read_Length_nt'


setwd("~/Exosomas/QC/readLen/Expresion_180913-190529/")
filenames <- list.files(pattern="*/*.tsv", recursive=TRUE) 
all_files <- lapply(filenames,function(x) { 
read.delim(file = x,
sep = '\t',
header = TRUE)
})
list3<- lapply(all_files, "[", c(1,5))
matrixA<-rbindlist(list3, fill=TRUE) 
readLen_RC<-matrixA %>% 
    group_by(Read.Length..nt.) %>% 
    dplyr::summarise(across(starts_with("x"), ~sum(., na.rm = TRUE)))

readLen_RC_D<-readLen_RC %>% dplyr::select(-contains("Brain"))
names(readLen_RC_D)=stringr::str_sub(names(readLen_RC_D),2)
names(readLen_RC_D)[names(readLen_RC_D) == 'ead.Length..nt.'] <- 'Read_Length_nt'


setwd("~/Exosomas/QC/readLen/Expresion_180919-190527/")
filenames <- list.files(pattern="*/*.tsv", recursive=TRUE) 
all_files <- lapply(filenames,function(x) { 
read.delim(file = x,
sep = '\t',
header = TRUE)
})
list3<- lapply(all_files, "[", c(1,5))
matrixA<-rbindlist(list3, fill=TRUE) 
readLen_RC<-matrixA %>%   
    group_by(Read.Length..nt.) %>% 
    dplyr::summarise(across(starts_with("x"), ~sum(., na.rm = TRUE)))

readLen_RC_E<-readLen_RC %>% dplyr::select(-contains("Brain"))
names(readLen_RC_E)=stringr::str_sub(names(readLen_RC_E),2)
names(readLen_RC_E)[names(readLen_RC_E) == 'ead.Length..nt.'] <- 'Read_Length_nt'


setwd("~/Exosomas/QC/readLen/Expresion_181003-190524/")
filenames <- list.files(pattern="*/*.tsv", recursive=TRUE) 
all_files <- lapply(filenames,function(x) { 
read.delim(file = x,
sep = '\t',
header = TRUE)
})
list3<- lapply(all_files, "[", c(1,5))
matrixA<-rbindlist(list3, fill=TRUE) 
readLen_RC<-matrixA %>%   
    group_by(Read.Length..nt.) %>% 
    dplyr::summarise(across(starts_with("x"), ~sum(., na.rm = TRUE)))

readLen_RC_F<-readLen_RC %>% dplyr::select(-contains("Brain"))
names(readLen_RC_F)=stringr::str_sub(names(readLen_RC_F),2)
names(readLen_RC_F)[names(readLen_RC_F) == 'ead.Length..nt.'] <- 'Read_Length_nt'


setwd("~/Exosomas/QC/readLen/Expresion_181004-190521/")
filenames <- list.files(pattern="*/*.tsv", recursive=TRUE) 
all_files <- lapply(filenames,function(x) { 
read.delim(file = x,
sep = '\t',
header = TRUE)
})
list3<- lapply(all_files, "[", c(1,5))
matrixA<-rbindlist(list3, fill=TRUE) 
readLen_RC<-matrixA %>%  
    group_by(Read.Length..nt.) %>% 
    dplyr::summarise(across(starts_with("x"), ~sum(., na.rm = TRUE)))

readLen_RC_G<-readLen_RC %>% dplyr::select(-contains("Brain"))
names(readLen_RC_G)=stringr::str_sub(names(readLen_RC_G),2)
names(readLen_RC_G)[names(readLen_RC_G) == 'ead.Length..nt.'] <- 'Read_Length_nt'

setwd("~/Exosomas/QC/readLen/Expresion_181023-190319/")
filenames <- list.files(pattern="*/*.tsv", recursive=TRUE) 
all_files <- lapply(filenames,function(x) { 
read.delim(file = x,
sep = '\t',
header = TRUE)
})
list3<- lapply(all_files, "[", c(1,5))
matrixA<-rbindlist(list3, fill=TRUE) 
readLen_RC<-matrixA %>%   
    group_by(Read.Length..nt.) %>% 
    dplyr::summarise(across(starts_with("x"), ~sum(., na.rm = TRUE)))

readLen_RC_H<-readLen_RC %>% dplyr::select(-contains("Brain"))
names(readLen_RC_H)=stringr::str_sub(names(readLen_RC_H),2)
names(readLen_RC_H)[names(readLen_RC_H) == 'ead.Length..nt.'] <- 'Read_Length_nt'

```



## Genome Distribution

Plotting the lenght distribution of reads by batch

```{r}
readLen_RC_A$A = rowMeans(readLen_RC_A[,c(2,36)])
readLen_RC_B$B = rowMeans(readLen_RC_B[,c(2,32)])
readLen_RC_C$C = rowMeans(readLen_RC_C[,c(2,36)])
readLen_RC_D$D = rowMeans(readLen_RC_D[,c(2,36)])
readLen_RC_E$E = rowMeans(readLen_RC_E[,c(2,35)])
readLen_RC_F$F = rowMeans(readLen_RC_F[,c(2,34)])
readLen_RC_G$G = rowMeans(readLen_RC_G[,c(2,35)])
readLen_RC_H$H = rowMeans(readLen_RC_H[,c(2,29)])

readLen<-data.frame(readLen_RC_A$Read_Length_nt,readLen_RC_A$A,readLen_RC_B$B,
                    readLen_RC_C$C,readLen_RC_D$D,readLen_RC_E$E,readLen_RC_F$F,readLen_RC_G$G,readLen_RC_H$H)
names(readLen)<-c("Length", "A","B","C","D","E","F","G","H")

readLen_long <-data.frame(reshape::melt(readLen, id.vars = "Length"))

colnames(readLen_long)<-c("Length", "Batch", "Value")

ggplot(readLen_long, aes(x=Length,y=Value,color=Batch)) +  
geom_line() + scale_color_igv() +  
  geom_vline(xintercept = 22,linetype="dashed")+ 
  annotate("text", x=23, y=0, label="22nt", angle=0, fontface = "bold") + 
  theme_bw() + 
labs(x = "Nucleotide length", y= "% Reads" , title="Genome distribution")+
  geom_line(size=1.2) +theme(axis.title = element_text(face="bold"), 
                            plot.title = element_text(size = 12, face = "bold")) 


```


# Multi dimensional Scaling

We observed a possible batch effect in the reads distribution, so we performed a MDS using DESeq2 package.



```{r echo=FALSE, message=FALSE}

# Format to expression matrix and clinical info table

setwd("/home/david_caceres/Exosomas/QC/Expresion/Matrices de Expresion/Expression matrix/")


matriz<-read.csv(file="matrixX.csv",header = TRUE, sep = ",", row.names=1)
matriz[is.na(matriz)] = 0
names(matriz)=stringr::str_sub(names(matriz),2) # Removing x from columns
names(matriz) <- gsub(x = names(matriz), pattern = "\\.", replacement = "-") # Adapting names

meta<-read.table(file="~/Exosomas/QC/Data/metadata.tsv", header = TRUE, sep = "\t")

cross<-read.csv(file="~/Exosomas/QC/Data/CrossSectional.minInfo.csv", header = TRUE, sep = ",")


# Format to metadata

meta2 <- strsplit(as.character(meta$sample),'_EXO_')
meta3<-do.call(rbind, meta2)
meta4<-cbind(meta, meta3)
colnames(meta4)= c("sample", "batch","batch_shortname", "OMICID", "smp" )

# Building a table with clinical info

cross2<-merge(cross, meta4, by="OMICID")
cross2[, 'batch_shortname'] <- as.factor(cross2[, 'batch_shortname'])
cross2$sample_shortname<-paste(cross2$batch_shortname, cross2$smp)
cross2$Diagnosis[is.na(cross2$Diagnosis)] = "Control"
cross2<-cross2[,-2]
Reads<-data.frame(tipos_exo$sample[1:266],tipos_exo$value[1:266])
colnames(Reads)<-c("sample","Reads")
cross2<-merge(cross2, Reads, by="sample")
Short_reads<-data.frame(short$sample[1:266],short$value[1:266])
colnames(Short_reads)<-c("sample","Short_reads")
cross2<-merge(cross2, Short_reads, by="sample")
Ultra_short_reads<-data.frame(ushort$sample[1:266],ushort$value[1:266])
colnames(Ultra_short_reads)<-c("sample","Ultra_short_reads")
cross2<-merge(cross2, Ultra_short_reads, by="sample")
bacteria <-contaminacion %>%
  filter(str_detect(variable, "bacteria"))
Bacteria<-data.frame(bacteria$sample,bacteria$value)
colnames(Bacteria)<-c("sample","Bacteria")
cross2<-merge(cross2, Bacteria, by="sample")
virus <-contaminacion %>%
  filter(str_detect(variable, "virus"))
Virus<-data.frame(virus$sample,virus$value)
colnames(Virus)<-c("sample","Virus")
cross2<-merge(cross2, Virus, by="sample")
reads_total<-read.table(file="~/Exosomas/QC/Data/Total_reads.tsv", header = TRUE, sep = "\t")
reads_total<-reads_total[- grep("Brain", reads_total$sample),]
reads_total<-data.frame(reads_total$sample,reads_total$reads)
colnames(reads_total)<-c("sample","reads_total")

cross2<-merge(cross2, reads_total, by="sample")

# Removing brain samples

matriz2<-matriz %>% dplyr::select(-contains("Brain"))
row.names(matriz2)<-matriz2$ame
matriz2<-matriz2 %>% dplyr::select(-contains("ame"))
matriz2<-matriz2 %>% mutate_if(is.integer,as.numeric)
matriz2<-matriz2 %>% mutate_if(is.character,as.numeric)

# Same order in samples for expression matrix and clinical info

data.table::setcolorder(matriz2,cross2$sample)

# Checking the same order

all.equal(colnames(matriz2), cross2$sample)


```

## DESeq2 object

```{r}
# Filtering by 50 counts per million and 20 samples

counts.CPM <- cpm(matriz2)
thresh <- counts.CPM > 50
keep <- rowSums(thresh) >= 20
counts.keep <- matriz2[keep,]

# Design matrix

design  <-  model.matrix(~0+cross2$Diagnosis)

# DEseq2 object

dds <- DESeqDataSetFromMatrix(counts.keep,
colData = cross2,
design = design)
```


## Multi dimenstional scaling plots

Plot of all the quality features to search for possible batch effect.

```{r echo=FALSE, warning=FALSE,message=FALSE}
# Multi dimensional scaling

vsd <- varianceStabilizingTransformation(dds,blind=TRUE)

mds <- cmdscale(dist(t(assay(vsd))))
cross2.mds <- merge(x = cross2,y = mds,by.x='sample',by.y='row.names')

colnames(cross2.mds)<-c("sample","OMICID","Age","Sex","Race","Diagnosis","Center","PatientID","Batch","batch_shortname","smp","sample_shortname","Reads","Short_reads","Ultra_short_reads","Bacteria","Virus","reads_total","V1","V2")

m1<-ggplot(cross2.mds, aes(x = V1, y = V2, color = batch_shortname)) +
    geom_point(size = 3) + coord_fixed() + ggtitle("MDS Batch") + stat_ellipse() + theme_bw() +theme(axis.title = element_text(face="bold"), plot.title = element_text(size = 12, face = "bold"))

m1<- m1 + scale_color_igv(name = "Batch")

m2<-ggplot(cross2.mds, aes(x = V1, y = V2, color = Sex)) +
    geom_point(size = 3) + coord_fixed() + ggtitle("MDS Sex") + stat_ellipse() + theme_bw()+theme(axis.title = element_text(face="bold"), plot.title = element_text(size = 12, face = "bold"))

m2<- m2 + scale_color_igv(name = "Sex")

m3<-ggplot(cross2.mds, aes(x = V1, y = V2, color = Diagnosis)) +
    geom_point(size = 3) + coord_fixed() + ggtitle("MDS Diagnosis") + stat_ellipse() + theme_bw()+theme(axis.title = element_text(face="bold"), plot.title = element_text(size = 12, face = "bold"))

m3<- m3 + scale_color_igv(name = "Diagnosis")

m5<-ggplot(cross2.mds, aes(x = V1, y = V2, color = Center)) +
    geom_point(size = 3) + coord_fixed() + ggtitle("MDS Center") + stat_ellipse() + theme_bw() + scale_color_igv()

m6<-ggplot(cross2.mds, aes(x = V1, y = V2, color = Age)) +
    geom_point(size = 3) + coord_fixed() + ggtitle("MDS Age") + theme_bw()+theme(axis.title = element_text(face="bold"), plot.title = element_text(size = 12, face = "bold"))

m6<- m6 + scale_color_gradient(low="white", high="red",name = "Age")

m7<-ggplot(cross2.mds, aes(x = V1, y = V2, color = Reads)) +
    geom_point(size = 3) + coord_fixed() + ggtitle("MDS Reads percent") + theme_bw() +theme(axis.title = element_text(face="bold"), plot.title = element_text(size = 12, face = "bold"))

m7<- m7 + scale_color_gradient(low="white", high="red",name = "Reads%")

m8<-ggplot(cross2.mds, aes(x = V1, y = V2, color = Short_reads)) +
    geom_point(size = 3) + coord_fixed() + ggtitle("MDS Short reads") + theme_bw() +theme(axis.title = element_text(face="bold"), plot.title = element_text(size = 12, face = "bold"))

m8<- m8 + scale_color_gradient(low="white", high="red",name = "Reads%")

m9<-ggplot(cross2.mds, aes(x = V1, y = V2, color = Ultra_short_reads)) +
    geom_point(size = 3) + coord_fixed() + ggtitle("MDS Ultra short reads") + theme_bw() +theme(axis.title = element_text(face="bold"), plot.title = element_text(size = 12, face = "bold"))

m9<- m9 + scale_color_gradient(low="white", high="red",name = "Reads%")

m10<-ggplot(cross2.mds, aes(x = V1, y = V2, color = Bacteria)) +
    geom_point(size = 3) + coord_fixed() + ggtitle("MDS Bacteria") + theme_bw() 

m10<- m10 + scale_color_gradient(low="white", high="dark blue",name = "Bacteria%")

m11<-ggplot(cross2.mds, aes(x = V1, y = V2, color = Virus)) +
    geom_point(size = 3) + coord_fixed() + ggtitle("MDS Virus") + theme_bw() 

m11<- m11 + scale_color_gradient(low="white", high="dark blue",name = "Virus%")

m12<-ggplot(cross2.mds, aes(x = V1, y = V2, color = reads_total, label=batch_shortname)) +
    geom_text(size = 5) + coord_fixed() + ggtitle("MDS Reads total") + theme_bw() +theme(axis.title = element_text(face="bold"), plot.title = element_text(size = 12, face = "bold"))

m12<- m12 + scale_color_gradient(low="white", high="red",name = "Reads total")

m1
m2
m3
m5
m6
m7
m8
m9
m10
m11
```


## Principal components correlation

Finally we performed a test to check the possible correlation between variables and find the real origin of the batch effect.


```{r fig.height=4, fig.width=6, message=FALSE, warning=FALSE}

cross3<-data.frame(cross2$Age,cross2$Gender,cross2$Diagnosis,cross2$batch_shortname, cross2$Reads,cross2$Short_reads,cross2$Ultra_short_reads,cross2$reads_total, cross2$Bacteria, cross2$Virus)
rownames(cross3)<-cross2$sample
colnames(cross3)<-c("Age","Gender","Diagnosis","Batch","Reads perc","Short Reads", "Ultra short reads", "Reads total", "Bacteria", "Virus")

all.equal(colnames(assay(vsd)), rownames(cross3))

cross3$Gender<-as.factor(cross3$Gender)
cross3$Age<-as.numeric(cross3$Age)
cross3$Diagnosis<-as.factor(cross3$Diagnosis)
names(cross3)[names(cross3) == 'Gender'] <- 'Sex'

pr_out <- prince(assay(vsd),cross3,top=20,center = TRUE)
prince.plot(pr_out,note = TRUE,Rsquared = TRUE)
confounding(cross3,method="chisq")


```
